---
title: "Synthetic US Survey Sample Generation"
author: "Armin Satzger"
date: "2022-11-20"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(psych)
library(httr)
library(jsonlite)
library(dplyr)
library(synthpop)
library(simstudy)
library(data.table)
library(synthpop)
library(ggplot2)
library(XML)
library(reshape2)
library(plyr)
library(stringr)
library(WriteXLS)


```


For the creation of our synthetic dataset, we have the choice between an empirically driven and a hypothesis driven approach. 

One way to implement a hypothesis-driven approach would, for example, be represented by using the *simstudy* R package for all survey attributes that allows one to parametrize variable distributions and relationships between variables. It does, however, not seem straightforward to find a suitable parametric distribution for some of the characteristics, like age, in particular, as heightened age-specific death factors are not easily represented by any common parametric distribution. As ample empirical evidence is available for at least four out of the five questions in the desired survey design, it, therefore, seems more appropriate to use publicly available data to inform the synthetic data generation process. 

Our goal for the synthetic survey sample is for it to be of size $n=2000$ and to be representative of the US population. As we want it to be representative of the overall population, the sampling method used does not play a major role and we could use both random choice as well as stratified sampling blocking on certain characteristics to curate our sample. 

For this reason, we may use publicly available US survey data as the starting point for our synthetic data generation exercise. One of the most high-quality large-scale, freely accessible and most recent sources of data on the US population is data of the US Census, more specifically the most recent, 2021 edition of the Public Use Microdata Sample (PUMS) as described on [the US Census website](https://www.census.gov/programs-surveys/acs/microdata.html).

## Obtain US Census data to base synthetic dataset on

After registering a new US Census API key, I download 2021 micro-level US Census data via the publicly available API [^1]:

[^1]: Please note that directly including access credentials such as API keys in the code constitutes a horrible security practice; common software engineering best practices would require storing credentials in a separate (.env) file but this is omitted here since the data is free to access and use.

```{r query-api}
# Define API link
link <- "https://api.census.gov/data/2021/acs/acs1/pumspr?get=SEX,AGEP,HINCP,SCHL&key=c0798cc737fff9005fc6501ac1a5c2575f76b0c6"

# Read in data at provided link as dataframe
test_data <- read.csv(link)
```


```{r prepare-data}
# Rename columns, remove auxiliary column X
clean_data <- dplyr::rename(subset(test_data, select=-c(X)), "sex"="X..SEX", age=AGEP, income=HINCP, education="SCHL.")

# Recode sex
# descriptions from https://api.census.gov/data/2021/acs/acs1/pums/variables/SEX.json
clean_data$sex <- as.factor(as.numeric(gsub("\\[", "", clean_data$sex)))
keys <- 1:2
values <- c("Male", "Female")
keysvals <- setNames(values, keys)
clean_data$sex <- recode(clean_data$sex, !!!keysvals)

# Recode income
# descriptions from https://api.census.gov/data/2021/acs/acs1/pumspr/variables/HINCP.json
clean_data$income <- na_if(clean_data$income, -60000) # sets values with missing value code as missing

# Recode education into six major categories: no high school, high school, vocational education, undergraduate, graduate, PhD
# descriptions from https://api.census.gov/data/2021/acs/acs1/pumspr/variables/SCHL.json
clean_data$education <- as.numeric(gsub("]", "", clean_data$education)) # remove ']' string pattern from string values
keys <- 0:24 # generate sequence vector with values from 0 to 24
values <- c(NA, rep("Did not complete high school", 15), "High school diploma", "Vocational education", rep("Some college", 3), "Bachelor's degree", rep("Master's degree", 2), "PhD") # create a vector with values corresponding to numeric keys
keysvals <- setNames(values, keys) # assign key-value pairs; create named vector
clean_data$education <- as.factor(recode(clean_data$education, !!!keysvals)) # recode edu values using mapping set up above
```

Please note that factor variables, in this case sex and education, are marked by an asterisk (*) next to the respective variable name. The number of data points for each question shows the largest number of missing values for the question about household income, followed by the question about educational attainment. This result might reflect hesitancy to share income information due to the social stigmata of unemployment/low social class for very low household incomes and potential jealousy for positive income outliers.
```{r summarize-data, include=FALSE, echo=TRUE}
psych::describe(clean_data, skew=FALSE, na.rm=TRUE)
```
```{r age-plots, include=FALSE, echo=TRUE}
ggplot(clean_data, aes(x=age, color=sex)) +
  geom_histogram(fill="white", alpha=0.5, position="identity")

pop_diagram <- clean_data

## cut the age variable into age groups with 5-year intervals
pop_diagram$AGEcut <- cut(pop_diagram$age, breaks = seq(0, 100, 5), right = FALSE) 
pop_diagram$Population <- 10 ## each sampled respondent represents 10 individuals
pop_diagram$Gender <- pop_diagram$sex

## aggregate the data by gender and age group
pop_diagram <- aggregate(Population ~ Gender + AGEcut, data = pop_diagram, FUN = sum)

## sort data by first by gender, then by age groups
pop_diagram <- with(pop_diagram, pop_diagram[order(Gender,AGEcut),])

## for simplicity, add the age group labels we used in pop_diagramcens above
#pop_diagram$Age <- rep(unique(popGHcens$Age)[1:20], 2)
vec1 <- as.character(seq(0,95,5))
vec2 <- as.character(seq(4,99,5))
vec3 <- str_c(vec1, "-", vec2)
pop_diagram$Age <- factor(rep(vec3, 2), levels=unique(rep(vec3,2)))

## only use the three variables age, gender and population from the pop_diagram data
pop_diagram <- pop_diagram[,c("Age","Gender","Population")]

## barplots for male populations goes to the left (thus negative sign)
pop_diagram$Population <- ifelse(pop_diagram$Gender == "Male", -1*pop_diagram$Population, pop_diagram$Population)

## pyramid charts are two barcharts with axes flipped
pyramid_census <- ggplot(pop_diagram, aes(x = Age, y = Population, fill = Gender)) + 
  geom_bar(data = subset(pop_diagram, Gender == "Female"), stat = "identity") +
  geom_bar(data = subset(pop_diagram, Gender == "Male"), stat = "identity") +
  scale_y_continuous(breaks = c(-10000, -5000, 0, 5000, 10000), labels = c(10000, 5000, 0, 5000, 10000)) +
  ylab("Sample observations") +
  ggtitle("Age Pyramid by Gender") +
  theme(plot.title = element_text(hjust = 0.5)) +
  coord_flip()

pyramid_census


```

## Create synthetic survey dataset with socioeconomic and demographic characteristics

```{r create-synthesized}
# Define seed in order to make replication possible
my.seed <- 19234834

# Generate synthetic data of sample size k=2000 based on adult-age (>=18 years of age) US Census sample
sds.default <- syn(filter(clean_data, age >= 18), seed=my.seed, k=2000)

# Assign created synthetic dataset dataframe to synthetic_data object
synthetic_data <- sds.default$syn
```

## Create life insurance dummy variable correlated with other attributes

Based on the information provided, life insurance has an overall prevalence of c. 60% across the adult population (previously filtered out underage respondents from the US Census data). Now, we also make a few additional assumptions that allow us to create the life insurance coverage attribute in the synthetic dataset: The older respondents are, the more likely it is they have dependents or other financial obligations that are financially dependent on the respondent's income, which implies a higher utility of life insurance and thus likely positive association between active life insurance and age. Additionally, women might be more inclined to take out life insurance as research has shown that women generally tend to be more risk averse than men.[^2]

[^2]: See, e.g., Eckel, C. C., & Grossman, P. J. (2008). Men, women and risk aversion: Experimental evidence. Handbook of experimental economics results, 1, 1061-1073.

```{r temp-var, include=FALSE, echo=FALSE}
synthetic_data$sex_numeric <- recode(synthetic_data$sex, "Male" = 0, "Female" = 1)
```

```{r create-life-insurance}
# Define distribution for life insurance status attribute
addDef <- defDataAdd(varname="insured", dist="binary", formula="0.5 + 0.002 * age + 0.02 * sex_numeric")

# Add column to dataset using specified distribution
synthetic_data <- setDF(addColumns(addDef, setDT(synthetic_data)))

# Recode insured variable to bool variable
synthetic_data$insured <- if_else(synthetic_data$insured == 1, TRUE, FALSE)
```

```{r temp-var-remove, include=FALSE, echo=FALSE}
synthetic_data <- subset(synthetic_data, select=-c(sex_numeric))
```


```{r save-data, include=FALSE}
save(synthetic_data, file="C:/Users/armin/Documents/synthetic_data.Rdata")
write.csv(synthetic_data, "C:/Users/armin/Documents/synthetic_data.csv")
```
